>

### 2020-09-22 manuscript/manuscript.Rmd
Various typos

---
title: >
  Modeling the impacts of park access on health outcomes: a utility-based 
  accessibility approach
runninghead: Macfarlane \emph{et al}.
author:
  - name: Gregory S. Macfarlane
    num: 1
  - name: Nico Boyd
    num: 2
  - name: John E. Taylor
    num: 3
  - name: Kari Watkins
    num: 3
address:
  - code: 1
    address: Civil and Environmental Engineering Department, Brigham Young University 
  - code: 2
    address: Fehr and Peers, Los Angeles, California
  - code: 3
    address: School of Civil and Environmental Engineering, Georgia Institute of Technology
corrauth: "Gregory Macfarlane, 430 Engineering Building, Provo UT 84663"
email: gregmacfarlane@gmail.com
abstract: >
  Recent research has underscored the potential for public green spaces to influence individual and societal health outcomes, but empirical measurements of such influences have yielded mixed results to date, with particular disagreement surrounding how access to parks ought to be defined while controlling for alternate explanations. In this paper we apply a comprehensive measure of park accessibility drawn from random utility choice theory, which avoids arbitrary assertions of proximity while incorporating potentially numerous amenities and attributes of both the parks and the population. We apply this utility-based accessibility measure to correlate Census tract-level obesity and physical activity rate estimates from the Centers for Disease Control and Prevention's 500 Cities project with tract-level American Community Survey socioeconomic data in New York City, paired with geographic open space data from New York City. Controlling for the socioeconomic variables and spatially correlated error terms, we show a positive and significant relationship between park access and physical activity rates. The data also suggest a negative relationship between park access and obesity rates, beyond what is expected through physical activity and socioeconomics. In doing so, this research contributes a more comprehensive modeling approach for measuring the impact of park access on health, and may improve our understanding of the role  parks and access to them can serve in furthering public health objectives.
keywords: Accessibility; Green Space
classoption:
  - shortAfour
output: 
  rticles::sage_article:
    keep_md: yes
    keep_tex: yes
    includes: 
      in_header: preamble.tex
      after_body: appendix.tex
bibliography: library.bib
bibliographystyle: sageh.bst
---

@@

To determine which spatial effects specification was appropriate, we estimated
SLX and SDEM models of physical activity participation rates as a function of
tract covariates and  accessibility. These two models are presented in Table 
\ref{tab:slx-sdem}. With only a few exceptions, the SLX and SDEM coefficient
estimates for corresponding covariates lie within the 95\% confidence intervals
of the other model. A Hausman-style test [@Pace2008] on the coefficients 
assumes a null hypothesis that the coefficients are equivalent; this test 
produces $p$-value of only $`r slx_sdem_htest[2]`$. Though this statistic is
well beneath a common hypothesis test threshold of $\alpha = 0.05$, we
elect retain the null hypothesis that the coefficients are equivalent, as 
the differences are at no point practically different from each other. The consequences
of rejecting the null hypothesis in this case would be to suggest a model with 
spatial dependence of the dependent variable; as stated above, we feel that such
a relationship is unreasonable. As a consequence of these analyses, we adopt the
SDEM specification moving forward as the spatial lag of the errors is 
significant and the overall log likelihood of the SDEM is higher than the SLX 
specification.

@@

\begin{acks}
This project was funded by the Speedwell Foundation. We are grateful to Rachel
Samuels and Neda Madi for helping to assemble the data.
\end{acks}

>>>>

---
title: >
  Modeling the impacts of park access on health outcomes: a utility-based 
  accessibility approach
runninghead: Macfarlane \emph{et al}.
author:
  - name: Gregory S. Macfarlane
    num: 1
  - name: Nico Boyd
    num: 2
  - name: John E. Taylor
    num: 3
  - name: Kari Watkins
    num: 3
address:
  - code: 1
    address: Civil and Environmental Engineering Department, Brigham Young University 
  - code: 2
    address: Fehr and Peers, Los Angeles, California
  - code: 3
    address: School of Civil and Environmental Engineering, Georgia Institute of Technology
corrauth: "Gregory Macfarlane, 430 Engineering Building, Provo UT 84663"
email: gregmacfarlane@byu.edu
abstract: >
  Recent research has underscored the potential for public green spaces to influence individual and societal health outcomes, but empirical measurements of such influences have yielded mixed results to date, with particular disagreement surrounding how access to parks ought to be defined while controlling for alternate explanations. In this paper we apply a comprehensive measure of park accessibility drawn from random utility choice theory, which avoids arbitrary assertions of proximity while incorporating potentially numerous amenities and attributes of both the parks and the population. We apply this utility-based accessibility measure to correlate Census tract-level obesity and physical activity rate estimates from the Centers for Disease Control and Prevention's 500 Cities project with tract-level American Community Survey socioeconomic data in New York City, paired with geographic open space data from New York City. Controlling for the socioeconomic variables and spatially correlated error terms, we show a positive and significant relationship between park access and physical activity rates. The data also suggest a negative relationship between park access and obesity rates, beyond what is expected through physical activity and socioeconomics. In doing so, this research contributes a more comprehensive modeling approach for measuring the impact of park access on health, and may improve our understanding of the role  parks and access to them can serve in furthering public health objectives.
keywords: Accessibility; Green Space
classoption:
  - shortAfour
output: 
  rticles::sage_article:
    keep_md: yes
    keep_tex: yes
    includes: 
      in_header: preamble.tex
      after_body: appendix.tex
bibliography: library.bib
bibliographystyle: sageh.bst
---

@@

To determine which spatial effects specification was appropriate, we estimated
SLX and SDEM models of physical activity participation rates as a function of
tract covariates and  accessibility. These two models are presented in Table 
\ref{tab:slx-sdem}. With only a few exceptions, the SLX and SDEM coefficient
estimates for corresponding covariates lie within the 95\% confidence intervals
of the other model. A Hausman-style test [@Pace2008] on the coefficients 
assumes a null hypothesis that the coefficients are equivalent; this test 
produces $p$-value of only $`r slx_sdem_htest[2]`$. Though this statistic is
well beneath a common hypothesis test threshold of $\alpha = 0.05$, we
elect to retain the null hypothesis that the coefficients are equivalent, as 
the differences are at no point practically different from each other. The consequences
of rejecting the null hypothesis in this case would be to suggest a model with 
spatial dependence of the dependent variable; as stated above, we feel that such
a relationship is unreasonable. As a consequence of these analyses, we adopt the
SDEM specification moving forward as the spatial lag of the errors is 
significant and the overall log likelihood of the SDEM is higher than the SLX 
specification.

@@

\begin{acks}
This project was funded by the Speedwell Foundation. We are grateful to Rachel
Samuels and Neda Madi for helping to assemble the data. We are also grateful for
the three anonymous reviewers, whose comments significantly strengthened the 
manuscript.
\end{acks}
